### 2025-01-16 ESC195 Lecture Notes/main.tex
jan 16

            % Fill the sector area between hyperbola, line, and x-axis
            \fill[blue, opacity=0.3, domain=0:1, variable=\t]
                (0, 0) -- plot ({cosh(\t)}, {sinh(\t)}) -- cycle;
        \end{tikzpicture}
    \end{center}
    \caption{Sector of the hyperbola}
    \centering
\end{figure}
\paragraph{Catenary} The hyperbolic cosine function describes the shape of a hanging chain or cable. The catenary is the curve formed by a chain hanging from two points. It is given by the equation:
\begin{equation}
    a \cosh\left(\frac{x}{a}\right),
\end{equation}
\begin{definition}[Hyperbolic Tangent]
    The hyperbolic tangent function is defined as:
    \begin{equation}
        \tanh(x) = \frac{\sinh(x)}{\cosh(x)} = \frac{e^x - e^{-x}}{e^x + e^{-x}}. 
    \end{equation}
\end{definition}
\paragraph{Derivative} The derivative of the hyperbolic tangent function resembles the derivative of the regular tangent function:
\begin{equation} \frac{d}{dx} \tanh(x) = \text{sech}^2(x). \end{equation}
\paragraph{Identities} The hyperbolic tangent function satisfies the following identities:
\begin{align}
    \tanh(x) &= \frac{\sinh(x)}{\cosh(x)}, \\
    \text{sech}^2(x) &= 1 - \tanh^2(x).
\end{align}
\paragraph{Secant, Cosecant, Cotangent} The hyperbolic secant, cosecant, and cotangent functions are defined similarly to the regular secant, cosecant, and cotangent functions. They are the \textbf{reciprocal} of the hyperbolic cosine, sine, and tangent functions, respectively.
\paragraph{Inverse Hyperbolic Functions} The inverse hyperbolic functions are defined as the inverse of the hyperbolic functions. They are denoted by $\sinh^{-1}(x)$, $\cosh^{-1}(x)$, $\tanh^{-1}(x)$, etc.
\begin{align}
    \text{arsinh } x &= \ln\left(x + \sqrt{x^2 + 1}\right) \\[10pt]
    \text{arcosh } x &= \ln\left(x + \sqrt{x^2 - 1}\right) \\[10pt]
    \text{artanh } x &= \frac{1}{2} \ln\left(\frac{1 + x}{1 - x}\right) \\[10pt]
    \text{arcsch } x &= \ln\left(\frac{1}{x} + \sqrt{\frac{1}{x^2} + 1}\right) \\[10pt]
    \text{arsech } x &= \ln\left(\frac{1}{x} + \sqrt{\frac{1}{x^2} - 1}\right) \\[10pt]
    \text{arcoth } x &= \frac{1}{2} \ln\left(\frac{x + 1}{x - 1}\right)
\end{align}
\appendix
\input{appendix.tex}
\end{document}

>>>>

            % Fill the sector area between hyperbola, line, and x-axis
            \fill[blue, opacity=0.3, domain=0:1, variable=\t]
                (0, 0) -- plot ({cosh(\t)}, {sinh(\t)}) -- cycle;
        \end{tikzpicture}
    \end{center}
    \caption{Sector of the hyperbola}
\end{figure}
\paragraph{Catenary} The hyperbolic cosine function describes the shape of a hanging chain or cable. The catenary is the curve formed by a chain hanging from two points. It is given by the equation:
    a \cosh\left(\frac{x}{a}\right),
\begin{definition}[Hyperbolic Tangent]
    The hyperbolic tangent function is defined as:
        \tanh(x) = \frac{\sinh(x)}{\cosh(x)} = \frac{e^x - e^{-x}}{e^x + e^{-x}}. 
\end{definition}
\paragraph{Derivative} The derivative of the hyperbolic tangent function resembles the derivative of the regular tangent function:
\begin{equation} \frac{d}{dx} \tanh(x) = \text{sech}^2(x). \end{equation}
\paragraph{Identities} The hyperbolic tangent function satisfies the following identities:
    \tanh(x) &= \frac{\sinh(x)}{\cosh(x)}, \\
    \text{sech}^2(x) &= 1 - \tanh^2(x).
\paragraph{Secant, Cosecant, Cotangent} The hyperbolic secant, cosecant, and cotangent functions are defined similarly to the regular secant, cosecant, and cotangent functions. They are the \textbf{reciprocal} of the hyperbolic cosine, sine, and tangent functions, respectively.
\paragraph{Inverse Hyperbolic Functions} The inverse hyperbolic functions are defined as the inverse of the hyperbolic functions. They are denoted by $\sinh^{-1}(x)$, $\cosh^{-1}(x)$, $\tanh^{-1}(x)$, etc.
    \text{arsinh } x &= \ln\left(x + \sqrt{x^2 + 1}\right) \\[10pt]
    \text{arcosh } x &= \ln\left(x + \sqrt{x^2 - 1}\right) \\[10pt]
    \text{artanh } x &= \frac{1}{2} \ln\left(\frac{1 + x}{1 - x}\right) \\[10pt]
    \text{arcsch } x &= \ln\left(\frac{1}{x} + \sqrt{\frac{1}{x^2} + 1}\right) \\[10pt]
    \text{arsech } x &= \ln\left(\frac{1}{x} + \sqrt{\frac{1}{x^2} - 1}\right) \\[10pt]
    \text{arcoth } x &= \frac{1}{2} \ln\left(\frac{x + 1}{x - 1}\right)
\section{Further Applications of Integrals}
\subsection{Arc Length}
\paragraph{Arc Length}
    Given the curve $C: y = f(x)$, the arc length of the curve between two points $a$ and $b$, where $y'(x)$ is continuous, we would like the compute the arc length of the curve:
    $$ L = \int_C ds$$
\paragraph{Formula} We can derive the formula for the arc length of a curve $y = f(x)$ between two points $a$ and $b$ as:
    \intertext{At $x_i$ and $x_{i+1}$, the length of the segment is:}
    \Delta s &\approx \sqrt{(x_{i+1} - x_i)^2 + (y_{i+1} - y_i)^2} \nonumber \\
            &= \sqrt{\Delta_i x^2 + \Delta_i y^2} \nonumber \\
    \intertext{We also have, by MVT:}
    \frac{\Delta y}{\Delta x} &= y'(c) \nonumber \\
    \Delta y &= y'(c) \Delta x \nonumber \\
    \Delta s &\approx \sqrt{\Delta x^2 + (y'(c) \Delta x)^2} \nonumber \\
            &= \sqrt{1 + y'(c)^2} \Delta x \nonumber \\
    \intertext{Taking a Riemann sum:}
    L &= \lim_{n \to \infty} \sum_{i=1}^n \sqrt{1 + y'(c_i)^2} \Delta x \nonumber \\
    &= \int_a^b \sqrt{1 + y'(x)^2} \, dx 
    Given $f(x) = x^\frac{3}{2}$ between $x = 0$ and $x = 44$, we would like to compute the arc length of the curve:
        L &= \int_0^{44} \sqrt{1 + \left(\frac{3}{2}x^{\frac{1}{2}}\right)^2} \, dx \\
        &= \int_0^{44} \sqrt{1 + \frac{9}{4}x} \, dx \\
        &= \frac{8}{27} \left(1 + \frac{9}{4}x\right)^{\frac{3}{2}} \Big|_0^{44} = 296
\subsection{Surface Area of Revolution}
\paragraph{Surface Area of Revolution} Given the curve $C: y = f(x)$, the surface area of the curve between two points $a$ and $b$, where $y'(x)$ is continuous, we would like the compute the surface area of the curve:
    $$ A = \int_C 2\pi y \, ds$$
\paragraph{Formula} Deriving the formula for the surface area of a curve $y = f(x)$ between two points $a$ and $b$ is similar to the arc length formula:
    \intertext{At $x_i$ and $x_{i+1}$, the length of the segment is:}
    \Delta s &\approx \sqrt{\Delta_i x^2 + \Delta_i y^2} \nonumber \\
    \intertext{The surface area of the segment is, by continuity:}
    A_i &= \pi (y_i + y_{i+1}) \Delta s = 2\pi y^\star \Delta s \quad 
    \intertext{As the difference between $y^\star$ and $y_i$ diminishes as $\Delta x \rightarrow 0$, taking a Riemann sum:}
    A &= \lim_{n \to \infty} \sum_{i=1}^n 2\pi y_i \sqrt{\Delta_i x^2 + \Delta_i y^2} \nonumber \\
    &= \int_a^b 2\pi y \sqrt{1 + y'(x)^2} \, dx
    Given $y = \sqrt{x}$ between $x = 0$ and $x = 1$, we would like to compute the surface area of the curve:
        A &= \int_0^1 2\pi \sqrt{x} \sqrt{1 + \frac{1}{4x}} \, dx \\
        &= \pi \int^1_0 \sqrt{4x + 1} \, dx \\
        \intertext{Let $u = 4x + 1$, then $du = 4dx$:}
        &= \frac{\pi}{4} \int^5_1 \sqrt{u} \, du \\
        &= \frac{\pi}{6}\left(5^\frac{3}{2} -1\right)
\subsection{Applications to Physics and Engineering}
\paragraph{Hydrostatic Pressure} The hydrostatic pressure and Force is given by the Archimedes' principle:
\begin{subequations}
        P &= \rho g \cdot d, \\
        F &= \rho g \cdot d \cdot A
\end{subequations}
\paragraph{Formula} Given a tank rectangular with width defined by $w = f(x)$, where $x$ is depth, we would like to compute the force exerted on the tank by the water:
    F &= \int_0^b \rho g x f(x) \cdot dx

\appendix
\input{appendix.tex}
\end{document}

### 2025-01-16 MAT185 Lecture Notes/main.tex
jan 16

\documentclass[11pt]{article}
\input{preamble}
\begin{document}
\thispagestyle{empty}
{\LARGE \bf MAT 185 Lecture Notes}\\
{\large Hei Shing Cheung}\\
Linear Algebra, Winter 2024 \hfill MAT 185\\
\section{Vector Space}
\begin{definition}[Vector Space in $\mathbb{R}$]
    A vectors space,${V}$, over $\mathbb{R}$ is a collection of \textbf{object} $\textbf{v} \in V$ s.t. the follow axioms are followed
    \begin{enumerate}
        \item \textbf{Addition Axioms}
        \begin{enumerate}
            \item \textbf{Closure Under Addition:} \\
            $\textbf{x}, \textbf{y} \in V \implies \textbf{x} + \textbf{y} \in V$

@@

            \item \textbf{Existence of Multiplicative Identity:} \\
            $\forall \, \textbf{x} \in V, \, 1 \textbf{x} = \textbf{x}$
        \end{enumerate}
    \end{enumerate}

@@

\paragraph{Results from above}

@@

\begin{theorem}[Commutativity of Addition]

@@

        \item $\textbf{x} + \textbf{y} \in W$
        \item $\alpha \textbf{x} \in W$
        \item $\textbf{0} \in W$

@@

@@

    Then, $\textbf{v}_1, \textbf{v}_2, \ldots, \textbf{v}_n$ are said to be linearly independent; otherwise, they are linearly dependent.
\end{definition}
\begin{definition}[Span]
    Let $V$ be a vector space and $\textbf{v}_1, \textbf{v}_2, \ldots, \textbf{v}_n \in V$, then:
    \begin{align*}
        \text{Span}(\textbf{v}_1, \textbf{v}_2, \ldots, \textbf{v}_n) = \left\{ \alpha_1 \textbf{v}_1 + \alpha_2 \textbf{v}_2 + \ldots + \alpha_n \textbf{v}_n \, | \, \alpha_1, \alpha_2, \ldots, \alpha_n \in \mathbb{R} \right\}
    \end{align*}
    is the set of all linear combinations of $\textbf{v}_1, \textbf{v}_2, \ldots, \textbf{v}_n$
\end{definition}

\end{document}

>>>>

\documentclass[11pt]{article}
\input{preamble}
\begin{document}
\thispagestyle{empty}
{\LARGE \bf MAT 185 Lecture Notes}\\
{\large Hei Shing Cheung}\\
Linear Algebra, Winter 2024 \hfill MAT 185\\
\section{Vector Space}
\begin{definition}[Vector Space in $\mathbb{R}$]
    This course concerns with real vector spaces. A vectors space, ${V}$, over $\mathbb{R}$ is a collection of \textbf{object} $\textbf{v} \in V$ s.t. the follow axioms are followed
    \begin{enumerate}
        \item \textbf{Addition Axioms}
        \begin{enumerate}
            \item \textbf{Closure Under Addition:} \\
            $\textbf{x}, \textbf{y} \in V \implies \textbf{x} + \textbf{y} \in V$

@@

            \item \textbf{Existence of Multiplicative Identity:} \\
            $\forall \, \textbf{x} \in V, \, \textbf{1} \textbf{x} = \textbf{x}$
        \end{enumerate}
    \end{enumerate}

@@

\paragraph{Additional Results from Above}

@@

\begin{theorem}[Addition is Commutative]

@@

        \item \textbf{Closure Under Addition:} \\
        $\textbf{x} + \textbf{y} \in W$
        \item \textbf{Closure Under Scalar Multiplication:} \\
        $\alpha \textbf{x} \in W$
        \item \text{Additive Identity:} \\
        $\textbf{0} \in W$

@@

\begin{definition}[Null Space]
    Let $V$ be a vector space and $A \in {^m\mathbb{R}^n}$\footnote{Same as $\mathbb{R}^{m\times n}$, The set of all $m \times n$ matrices with real entries}, then:
        \text{null}(A) = \left\{ \textbf{x} \in {^n\mathbb{R}} \, | \, A\textbf{x} = \textbf{0} \right\}
    is the null space of $A$, otherwise known as the \textbf{kernel} of $A$ or the solution space of $A\textbf{x} = \textbf{0}$
    \paragraph{} We can use the Subspace Test I to show that the null space of a matrix is a subspace of $^n\mathbb{R}$.
            \item \textbf{Existence of Additive Identity:} \\
            The zero vector is in the null space of $A$ as the trivial solution to the equation $A\textbf{x} = \textbf{0}$
            \item \textbf{Closure Under Addition:} \\
            Let $\textbf{x}, \textbf{y} \in \text{null}(A)$, then $A(\textbf{x} + \textbf{y}) = A\textbf{x} + A\textbf{y} = \textbf{0} + \textbf{0} = \textbf{0}$. Thus, $\textbf{x} + \textbf{y} \in \text{null}(A)$. This holds since $^n\mathbb{R}$ is a vector space.
            \item \textbf{Closure Under Scalar Multiplication:} \\
            Let $\textbf{x} \in \text{null}(A)$ and $\alpha \in \mathbb{R}$, then $A(\alpha \textbf{x}) = \alpha A\textbf{x} = \alpha \textbf{0} = \textbf{0}$. Thus, $\alpha \textbf{x} \in \text{null}(A)$. This holds since $^n\mathbb{R}^m$ is a vector space under usual addition and scalar multiplication.
\begin{theorem}[Subspace Test, II]
    Let $V$ be a vector space over $\mathbb{R}$ and $W \subseteq V$ with the usual addition and scalar multiplication, then $W$ is a subspace of $V$ iff for all $\textbf{x}, \textbf{y} \in W$ and $\alpha, \beta \in \mathbb{R}$:
        \item $\alpha \textbf{x} + \beta \textbf{y} \in W$
    ($\Leftarrow$) Assume the condition is satisfied, then $W$ is a vector space. Thus, $W$ is a subspace of $V$.

@@

    Then, $\textbf{v}_1, \textbf{v}_2, \ldots, \textbf{v}_n$ are said to be linearly independent; otherwise, they are linearly dependent.
\end{definition}
\begin{definition}[Span]
    Let $V$ be a vector space and $\textbf{v}_1, \textbf{v}_2, \ldots, \textbf{v}_n \in V$, then:
    \begin{align*}
        \text{Span}(\textbf{v}_1, \textbf{v}_2, \ldots, \textbf{v}_n) = \left\{ \alpha_1 \textbf{v}_1 + \alpha_2 \textbf{v}_2 + \ldots + \alpha_n \textbf{v}_n \, | \, \alpha_1, \alpha_2, \ldots, \alpha_n \in \mathbb{R} \right\}
    \end{align*}
    is the set of all linear combinations of $\textbf{v}_1, \textbf{v}_2, \ldots, \textbf{v}_n$
\end{definition}
\end{document}

### 2025-01-16 MSE160 Lecture Notes/main.tex
Add Potential Energy-Separation Curve and update lecture notes
jan 16

    \end{axis}
    \end{tikzpicture}
    \caption{Potential Energy-Separation Curve}
    \label{fig:potential-energy}
\end{figure}
\paragraph{Coefficient of Thermal Expansion} The coefficient of thermal expansion, $\alpha$, is the fractional change in length per degree change in temperature. 
\paragraph{Depth of Potential Energy Curve} The deeper the potential energy curve, the higher the melting temperature and more symetric the curve near $E_0$. This would give the follwoing three properties:
\begin{enumerate}
    \item \textbf{Higher Melting Temperature} The higher the melting temperature, the deeper the potential energy curve.
    \item \textbf{Higher Elastic Modulus} The steeper the slope of the force-seperation curve at $r_0$, the higher the elastic modulus.
    \item \textbf{Lower Coefficient of Thermal Expansion} The more symetric the potential energy curve near $E_0$, the lower the coefficient of thermal expansion.
\end{enumerate}
\subsection{Shear and Tensile Stress}
\subsubsection{Shear}
\paragraph{Shear Stress} Shear stress is the force per unit area acting parallel to the surface. It is given by:
\begin{equation}
    \tau = \frac{F}{A_0}
\end{equation}
\paragraph{Shear Strain} Shear strain is the change in angle between two lines originally perpendicular to each other. It is given by:
\begin{equation}
    \gamma = \frac{\Delta l}{l_0} \approx \tan \theta \approx \theta = \frac{\pi}{2} - \phi
\end{equation}
\paragraph{Shear Modulus} The shear modulus, $G$, is the ratio of shear stress to shear strain. It is given by:
\begin{equation}
    \tau = G \gamma
\end{equation}
\paragraph{Relationship between Shear and Tensile Modulus} The shear modulus is related to the tensile modulus by the following equation:
\begin{equation}
    G = \frac{E}{2(1+\nu)}
\end{equation}
where $\nu$ is the Poisson's ratio.
\paragraph{Poisson's Ratio} Poisson's ratio, $\nu$, is the ratio of lateral strain to axial strain. It is given by:
\begin{equation}
    \nu = -\frac{\epsilon_{\text{lat}}}{\epsilon_{\text{axial}}}
\end{equation}
\includepdf[pages=-, angle=90]{EquationSheet.pdf}
\end{document}

>>>>

    \end{axis}
    \end{tikzpicture}
    \caption{Potential Energy-Separation Curve}
    \label{fig:potential-energy}
\end{figure}
\paragraph{Coefficient of Thermal Expansion} The coefficient of thermal expansion, $\alpha$, is the fractional change in length per degree change in temperature. 
\paragraph{Depth of Potential Energy Curve} The deeper the potential energy curve, the higher the melting temperature and more symetric the curve near $E_0$. This would give the follwoing three properties:
\begin{enumerate}
    \item \textbf{Higher Melting Temperature} The higher the melting temperature, the deeper the potential energy curve.
    \item \textbf{Higher Elastic Modulus} The steeper the slope of the force-seperation curve at $r_0$, the higher the elastic modulus.
    \item \textbf{Lower Coefficient of Thermal Expansion} The more symetric the potential energy curve near $E_0$, the lower the coefficient of thermal expansion.
\end{enumerate}
\subsection{Shear and Tensile Stress}
\subsubsection{Shear}
\paragraph{Shear Stress} Shear stress is the force per unit area acting parallel to the surface. It is given by:
\begin{equation}
    \tau = \frac{F}{A_0}
\end{equation}
\paragraph{Shear Strain} Shear strain is the change in angle between two lines originally perpendicular to each other. It is given by:
\begin{equation}
    \gamma = \frac{\Delta l}{l_0} \approx \tan \theta \approx \theta = \frac{\pi}{2} - \phi
\end{equation}
\paragraph{Shear Modulus} The shear modulus, $G$, is the ratio of shear stress to shear strain. It is given by:
\begin{equation}
    \tau = G \gamma
\end{equation}
\paragraph{Relationship between Shear and Tensile Modulus} The shear modulus is related to the tensile modulus by the following equation:
\begin{equation}
    G = \frac{E}{2(1+\nu)}
\end{equation}
where $\nu$ is the Poisson's ratio.
\paragraph{Poisson's Ratio} Poisson's ratio, $\nu$, is the ratio of lateral strain to axial strain. It is given by:
\begin{equation}
    \nu = -\frac{\epsilon_{\text{lat}}}{\epsilon_{\text{axial}}}
\end{equation}
\subsection{Testing}
\begin{definition}[Gauge Length]
The gauge length, $l_0$, is the length of the sample over which the strain is measured.
\end{definition}
\begin{definition}[Reduced Section]
The reduced section is the part of the sample where the cross-sectional area is reduced to a smaller value.
\end{definition}
\paragraph{} Gauge length is always no longer than the reduced section. The reduced section is where the sample will likely break.
\paragraph{Testing Ceremics} In relation to tensile testing, ceramics have the following properties:
\begin{itemize}
    \item \textbf{Brittle} Ceramics are brittle and will break suddenly.
    \item \textbf{High Strength} Ceramics have high strength and thus difficult to machine the sample.
    \item \textbf{Sample Alignment} The sample must be aligned properly to test for pure tension. Unlike metals and polymers, which are self-aligning.
    \item \textbf{Fracture} Ceramics will fracture while still off-axis. Hence, there would be a large shear component. 
\end{itemize}
\paragraph{} Thus, we often approxiate tensile behaviour with a point loading on a horizontal beam, with two point support (3 point bending test). Peak stress is given by:
\begin{equation}
    \sigma_{\text{peak}} = \frac{3FL}{2bd^2}
\end{equation}
, where:
\begin{itemize}
    \item $L$ (span) is the distance between the two supports.
    \item $b$ is the width of the sample
    \item $d$ is the thickness/depth of the sample
\end{itemize}
\section{Seleection of Materials}
\begin{example}[Aircraft Wing Spar]
    The aircraft wing spar is beam (loaded in bending) that supports the wing. The spar is made of a material with the objective of minimize mass under the following constraints:
    \begin{itemize}
        \item \textbf{Deflection} There is a maximum allowable deflection of the wing.
        \item There is more..., but for this example, we will only consider the deflection.
    \end{itemize}
    The material selection solve for a \textbf{light stiff beam}.

    \paragraph{Mass} The mass of the beam is given by:
    $$ m = \rho V = \rho A L $$
    \paragraph{Deflection} The deflection of the beam is given by:
    \begin{align*} 
        \delta &= \frac{FL^3}{48EI} 
        \intertext{For a beam with a rectangular cross-section, we have:}
        \delta &= \frac{FL^3}{48E} \cdot \frac{12}{bh^3} = \frac{FL^3}{4Ebh^3}  
        \intertext{We can set $b$ proportional to $h$:}
        \delta &= \frac{FL^3}{cE} \cdot \frac{1}{A^2},\quad \text{for some constant $c$} 
        \intertext{We can then isolate for $A$, the free variable, and minimize the mass via the objective equation $m = \rho A L$:}
        A &= \sqrt{\frac{FL^3}{cE\delta}} \\
        m &= \rho L \sqrt{\frac{FL^3}{cE\delta}} = \rho L \sqrt{\frac{FL^3}{cE\delta}}
        \intertext{Arrange into the form (functional)(geometric)(material):}
        m &= \left(\frac{F}{c\delta}\right)^\frac{1}{2} \cdot \left(L^\frac{5}{2}\right)\left(\frac{\rho}{E^\frac{1}{2}}\right)
    \end{align*}
\end{example}
\paragraph{Material Peformance Index} The material performance index is given by:
    \begin{equation}
        \text{Material Performance Index (MSI)} = \frac{E^\frac{1}{2}}{\rho}
\end{equation}
\paragraph{MPI Graph} We plot $\log E$ against $\log \rho$ to get the MPI graph. 
\appendix
\includepdf[pages=-, angle=90]{EquationSheet.pdf}
\end{document}
